### 2021-02-12 ch05_paper_ecs/paper_ecs.tex
Optimized introduction of chapter 05

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous chapter begs the question about the interpretation and credibility of
these results. Since the \acs{CMIP}6 models provide crucial input for the
upcoming international climate assessment of the \ac{IPCC} (\acs{AR}6), this is
not only important for the scientific community, but also directly affects
society and policymakers. An important question in this context could be the
following: Is it reasonable to adapt the assessed \ac{ECS} range from \acs{AR}5
($\range{1.5}{4.5} \unit{K}$) in \acs{AR}6 based on the \acs{CMIP}6 model range
of $\range{1.8}{5.6} \unit{K}$? An argument for this is that due to the many
new and improved processes included to the models, the \acs{CMIP}6 ensemble is
arguably more realistic than its predecessor generations. However, even the
\acs{CMIP}6 models are far from fully simulating all aspects and processes of
the full vastly complex Earth system. This could introduce new biases which
have not been present in earlier \ac{CMIP} models. For example, the apparent
increase in climate sensitivity in the \acs{CMIP}6 models might be wrongly
caused by missing processes that cancel out the effects of the newly included
processes. Thus, it is vital to take other sources of information into account
before answering the question raised above.

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
system can be utilized within the emergent constraints framework to assess the
accuracy of climate model projections and reduce associated uncertainties. For
\ac{ECS}, a large variety of different emergent constraints has been developed
over the years, which we now evaluate on the \acs{CMIP}6 ensemble in order to
assess whether they still hold for these new climate models. In total we assess
eleven emergent constraints on \ac{ECS} (see
\cref{tab:05:overview_emergent_constraints}), out of which the most are related
to cloud feedbacks since these constitute the most important source of
uncertainty for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see also
  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, the
\acs{CMIP}6 models offer an exciting possibility to perform out-of-sample
testing using the previously unseen data.

@@

\begin{table}[p]
  \centering
  \hbadness=3500
  \begin{tabular}{p{0.07\columnwidth} p{0.24\columnwidth} p{0.29\columnwidth}
      p{0.29\columnwidth}}
    \toprule
    Label & Reference & Description of \xaxis{} & Variables and corresponding
    observations \\
    \midrule
    BRI & \textcite{Brient2016} & Sensitivity of shortwave cloud albedo
    to changes in \acf{SST} [$\si{\pctK{}}$] & \makecell{\tabitem\emph{hur}
      (ERA-Interim) \\ \tabitem \emph{rsdt} (CERES-EBAF) \\ \tabitem
      \emph{rsut} (CERES-EBAF) \\ \tabitem \emph{rsutcs} (CERES-EBAF) \\
      \tabitem \emph{ts} (HadISST)} \\
    \midrule
    COX & \textcite{Cox2018} & $\Psi$ (temperature variability metric)
    [$\si{K}$] &
    \tabitem \emph{tas} (HadCRUT4) \\
    \midrule
    LIP & \textcite{Lipat2017} & Extent of Southern Hemisphere Hadley cell
    [$\si{\degree}$] & \tabitem \emph{va} (ERA-Interim) \\
    \midrule
    SHD & \textcite{Sherwood2014} & $D$-index (large-scale lower-tropospheric
    mixing) [$\si{1}$] & \tabitem \emph{wap} (ERA-Interim) \\
    \midrule
    SHL & \textcite{Sherwood2014} & \acs{LTMI} (\acl{LTMI}) [$\si{1}$] &
    \makecell{\tabitem \emph{hur} (ERA-Interim) \\ \tabitem \emph{ta}
      (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SHS & \textcite{Sherwood2014} & $S$-index (small-scale lower-tropospheric
    mixing) [$\si{1}$] & \makecell{\tabitem \emph{hur} (ERA-Interim) \\
      \tabitem \emph{ta} (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SU & \textcite{Su2014} & Error in vertical profile of relative humidity
    [$\si{1}$] & \tabitem \emph{hur} (AIRS, MLS-Aura) \\
    \midrule
    TIH & \textcite{Tian2015} & Tropical mid-tropospheric humidity asymmetry
    index from AIRS [$\si{\%}$] & \tabitem \emph{hus} (AIRS) \\
    \midrule
    TII & \textcite{Tian2015} & Southern \acs{ITCZ} index from GPCP
    [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
    \midrule
    VOL* & \textcite{Volodin2008} & Difference between tropical and southern
    midlatitudinal cloud fraction [$\si{\%}$] & \tabitem \emph{clt} (ISCCP
    D-2) \\
    \midrule
    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} cloud
    fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
    \makecell{\tabitem \emph{cl} (Cloudsat/CALIPSO) \\ \tabitem \emph{tos}
      (AMSRE \ac{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \bottomrule
  \end{tabular}
  \caption{Overview of the eleven emergent constraints on the \acf{ECS}
    evaluated in this chapter. Detailed descriptions of the variables used to
    calculate the \xaxis{} values of the emergent constraints and the
    references for the corresponding observations (fourth column) are given in
    \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
    For emergent constraints marked with an asterisk (*), the observational
    value of the original publication is used to calculate the observational
    constraint. For all others, the observational value is computed with the
    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.}
  \label{tab:05:overview_emergent_constraints}
\end{table}

>>>>

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous chapter begs the question about the interpretation and credibility of
these results. Since the \acs{CMIP}6 models provide crucial input for the
upcoming international climate assessment of the \ac{IPCC} (\acs{AR}6), this is
not only important for the scientific community, but also directly affects
society and policymakers. Due to the many new and improved processes included
to the models, the \acs{CMIP}6 ensemble is arguably more realistic than its
predecessor generations. However, does this automatically make results derived
from \acs{CMIP}6 models more trustworthy than those derived from earlier
\ac{CMIP} models? In general, the answer to this question is \enquote{no}. Even
the \acs{CMIP}6 models are far from fully simulating all aspects and processes
of the full vastly complex Earth system, which could introduce new biases which
have not been present in earlier \ac{CMIP} models. For example, the apparent
increase in climate sensitivity in the \acs{CMIP}6 models might be wrongly
caused by missing processes that cancel out the effects of the newly included
processes. Thus, to assess the credibility of \ac{ESM} results it is vital to
take other sources of information into account.

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
system can be utilized within the so-called \emph{emergent constraints}
framework to assess the accuracy of climate model projections and reduce
associated uncertainties. For \ac{ECS}, a large variety of different emergent
constraints has been developed over the years, which we now evaluate on the
\acs{CMIP}6 ensemble in order to assess whether they still hold for these new
climate models, \ie{} whether they have skill to reduce uncertainties in
\ac{ECS} in the \acs{CMIP}6 ensemble. In total we assess eleven emergent
constraints on \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out
of which the most are related to cloud feedbacks since these constitute the
most important source of uncertainty for \ac{ECS} \commentcite{Boucher2013,
  Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
evaluated emergent constraints have been derived on the \acs{CMIP}3 or
\acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
perform out-of-sample testing using the previously unseen data.

\begin{table}[!t]
  \centering
  \hbadness=3500
  \begin{tabular}{p{0.07\columnwidth} p{0.24\columnwidth} p{0.29\columnwidth}
      p{0.29\columnwidth}}
    \toprule
    Label & Reference & Description of \xaxis{} & Variables and corresponding
    observations \\
    \midrule
    BRI & \textcite{Brient2016} & Sensitivity of shortwave cloud albedo
    to changes in \acf{SST} [$\si{\pctK{}}$] & \makecell{\tabitem\emph{hur}
      (ERA-Interim) \\ \tabitem \emph{rsdt} (CERES-EBAF) \\ \tabitem
      \emph{rsut} (CERES-EBAF) \\ \tabitem \emph{rsutcs} (CERES-EBAF) \\
      \tabitem \emph{ts} (HadISST)} \\
    \midrule
    COX & \textcite{Cox2018} & $\Psi$ (temperature variability metric)
    [$\si{K}$] &
    \tabitem \emph{tas} (HadCRUT4) \\
    \midrule
    LIP & \textcite{Lipat2017} & Extent of Southern Hemisphere Hadley cell
    [$\si{\degree}$] & \tabitem \emph{va} (ERA-Interim) \\
    \midrule
    SHD & \textcite{Sherwood2014} & $D$-index (large-scale lower-tropospheric
    mixing) [$\si{1}$] & \tabitem \emph{wap} (ERA-Interim) \\
    \midrule
    SHL & \textcite{Sherwood2014} & \acs{LTMI} (\acl{LTMI}) [$\si{1}$] &
    \makecell{\tabitem \emph{hur} (ERA-Interim) \\ \tabitem \emph{ta}
      (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SHS & \textcite{Sherwood2014} & $S$-index (small-scale lower-tropospheric
    mixing) [$\si{1}$] & \makecell{\tabitem \emph{hur} (ERA-Interim) \\
      \tabitem \emph{ta} (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \midrule
    SU & \textcite{Su2014} & Error in vertical profile of relative humidity
    [$\si{1}$] & \tabitem \emph{hur} (AIRS, MLS-Aura) \\
    \midrule
    TIH & \textcite{Tian2015} & Tropical mid-tropospheric humidity asymmetry
    index from AIRS [$\si{\%}$] & \tabitem \emph{hus} (AIRS) \\
    \midrule
    TII & \textcite{Tian2015} & Southern \acs{ITCZ} index from GPCP
    [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
    \midrule
    VOL* & \textcite{Volodin2008} & Difference between tropical and southern
    midlatitudinal cloud fraction [$\si{\%}$] & \tabitem \emph{clt} (ISCCP
    D-2) \\
    \midrule
    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} cloud
    fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
    \makecell{\tabitem \emph{cl} (Cloudsat/CALIPSO) \\ \tabitem \emph{tos}
      (AMSRE \ac{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
    \bottomrule
  \end{tabular}
  \caption{Overview of the eleven emergent constraints on the \acf{ECS}
    evaluated in this chapter. Detailed descriptions of the variables used to
    calculate the \xaxis{} values of the emergent constraints and the
    references for the corresponding observations (fourth column) are given in
    \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
    For emergent constraints marked with an asterisk (*), the observational
    value of the original publication is used to calculate the observational
    constraint. For all others, the observational value is computed with the
    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.}
  \label{tab:05:overview_emergent_constraints}
\end{table}

@@
